### 2020-04-19 serial.tex
more more more

Here, we provide an answer to this paradox, by showing that the relevant interval for the renewal framework is what is called the ``forward'' interval, and that the forward serial (but not generation) interval is affected by the rate of growth $r$ during the exponential-growth phase.
We develop a new framework for understanding serial intervals and show that the forward serial-interval distribution gives the correct value of $\mathcal R$.
%\jd{Come back to this. Not just over time, right?} \swp{What do you mean? Everything is about time here... you haven't read the results section so maybe that's why?}
However, using inaccurately defined serial intervals or failing to account for changes in the observed serial-interval distributions over the course of an epidemic can bias the estimate of $\mathcal R$.
We apply our framework to serial intervals of COVID-19 and lay out several principles to consider in using information about serial intervals and other epidemiological time delays in the analysis of the ongoing pandemic.

@@

In order to define forward and backward serial interval distributions, we begin by writing the total number of serial intervals between time $\pi$ (when the infector developed symptoms) and $\delta$ (when the infectee developed symptoms), which can be expressed as an integral over all of the possibilities for the infection times of the infector $\alpha_1$ and the infectee $\alpha_2$:
\begin{equation}
T(\pi,\delta) = \int_{-\infty}^{\pi} \int_{\alpha_1}^{\delta} \mathcal R_c (\alpha_1) i(\alpha_1) h_{\alpha_1}(\pi-\alpha_1, \alpha_2 - \alpha_1) k_{\alpha_2}(\delta - \alpha_2) \mathrm{d}\alpha_2\,\mathrm{d}\alpha_1,
\end{equation}
where $\mathcal R_c(\alpha_1)$ is the case reproduction number (i.e., the average number of secondary cases caused by a primary case infected at time $\alpha_1$, \cite{fraser2007estimating}), $i(\alpha_1)$ is incidence, $h_{\alpha_1}(\pi-\alpha_1, \alpha_2 - \alpha_1)$ is the joint probability distribution describing the forward incubation period $\pi-\alpha_1$ and forward generation interval $\alpha_2 - \alpha_1$ of an individual infected at time $\alpha_1$ (in this case, the infector), and $k_{\alpha_2}(\delta-\alpha_2)$ describes the marginal probability distribution of $h_{\alpha_2}$ describing the forward incubation period $\delta-\alpha_2$ of an individual infected at time $\alpha_2$ (in this case, the infectee). 
Then, the forward serial-interval distribution is proportional to $T(\pi, \pi+\tau)$. Substituting $x=\pi-\alpha_1$ and $\sigma=\alpha_2-\alpha_1$, we have:
\begin{equation}
f_\pi(\tau) \propto \int_{0}^{\infty} \int_{0}^{x+\tau} \mathcal R_c (\pi-x) i(\pi-x) h_{\pi-x}(x, \sigma) k_{\pi-x+\sigma}(x-\sigma+\tau) \mathrm{d}\sigma\,\mathrm{d}x
\end{equation}
Likewise, the backward serial-interval distribution is proportional to $T(\delta-\tau, \delta)$. 
This time, we substitute $x=\delta-\alpha_2$ and $\sigma=\alpha_2-\alpha_1$ to get:
\begin{equation}
b_\delta(\tau) \propto \int_{0}^{\infty} \int_{\max(0, \tau-x)}^{\infty} \mathcal R_c (\delta-x-\sigma) i(\delta-x-\sigma) h_{\delta-x-\sigma}(x+\sigma-\tau, \sigma) k_{\delta-x}(x) \mathrm{d}\sigma\,\mathrm{d}x.
\end{equation}

%Hereafter, we assume that the forward incubation period distributions does not vary across cohorts over the course of an epidemic, as they represent the natural history of a disease, and use $k$ without subscript instead:
%\begin{equation}
%k(x) = \int_0^\infty h_\pi(x, \sigma) \mathrm{d}\sigma.
%\end{equation}

@@

We validate the theory by applying it to a specific example of an epidemic model. 
We model disease spread with a renewal-equation model \citep{heesterbeek1996concept, diekmann2000mathematical, roberts2004modelling, aldis2005integral, roberts2007model, champredon2018equivalence}.
Ignoring births and deaths, changes in the proportion of susceptible individuals $S(t)$ and incidence of infection $i(t)$ can be written as:
\begin{equation}
\begin{aligned}
\frac{\mathrm{d}S}{\mathrm{d}t} &= - i(t)\\
i(t) &= \mathcal R_0 S(t) \int_0^\infty i(t-\tau) g(\tau) \mathrm{d}\tau,
\end{aligned}
\label{eq:renewal}
\end{equation}
where $\mathcal R_0$ is the basic reproduction number, and $g(\tau)$ is the intrinsic generation-interval distribution (i.e., the forward generation-interval distribution of a primary case in a fully susceptible population; \cite{champredon2015intrinsic}).
Then, the forward generation-interval for a primary cohort $\ell$ follows \citep{champredon2015intrinsic}:
\begin{equation}
g_\ell (\tau) \propto g(\tau) S(\ell + \tau),
\end{equation}
which allows us to separate the joint probability distribution $h_\ell$ of the forward incubation period and the forward generation-interval distribution as a product of the proportion of susceptible individuals $S$ and the joint probability distribution $h$ of the forward incubation period and the intrinsic generation intervals:
\begin{equation}
h_\ell (x_0, \tau) \propto h(x_0, \tau) S(\ell + \tau),
\end{equation}
which satisfies the following:
\begin{equation}
g(\tau) = \int_0^\infty h(x_0, \tau) \mathrm{d}x_0.
\end{equation}
Finally, the cohort reproduction for this model is defined as follows:
\begin{equation}
\mathcal R_c(t) = \mathcal R_0 \int_0^\infty g(\tau) S(t+\tau) \mathrm{d} \tau.
\end{equation}

@@

During the initial phase of an epidemic, the proprotion susceptible remains constant ($S(t) = S(0)$) and incidence of infection grows exponentially: $i(t)=i_0\exp(rt)$.
Then, we can estimate the reproduction number from the exponential growth rate $r$ via the Euler-Lotka equation:
\begin{equation}
\frac{1}{\mathcal R} = \int_0^\infty \exp(-r\tau) g(\tau) \mathrm{d} \tau.
\end{equation}
Like forward generation-interval distributions, 
forward serial-interval distributions describe the renewal process of symptomatic cases.
Therefore, the forward serial-interval distribution $f_{\textrm{\tiny exp}}(\tau)$ during the exponential growth phase provides the identical $r$--$\mathcal R$ link as the intrinsic generation-interval distribution:
\begin{equation}
\frac{1}{\mathcal R} = \int_{-\infty}^\infty \exp(-r\tau) f_{\textrm{\tiny exp}}(\tau) \mathrm{d} \tau,
\end{equation}
where the forward serial-interval distribution during the exponential growth phase is defined as:
\begin{equation}
f_{\textrm{\tiny exp}}(\tau) \propto \int_{0}^\infty \int_{0}^\infty \exp( -r x_0) h(x_0, \sigma) k(\tau-\sigma+x_0) \mathrm{d} x_0\, \mathrm{d}\sigma.
\end{equation}
In Appendix, we provide a mathematical proof that this relationship holds.

We note that the forward serial-interval distribution depends on the exponential growth rate $r$.
When the epidemic grows fast (high $r$), we expect the backward incubation period to be short, and therefore, the forward serial-interval distribution will generally have a larger mean than the intrinsic generation-interval distribution.
The Susceptible-Exposed-Infected-Recovered model, which assumes that incubation and exposed periods are equivalent, is a special case where the conditional forward generation-interval distribution cancels out with the backward generation-interval distribution exactly because (i) infected individuals can only transmit after symptom onset and (ii) the time between symptom onset to infection is independent of the incubation period of an infector;
in this case, the forward serial- and generation-intervals have the same distributions during the exponential growth phase.

@@

We use a simulation-based approach to compare the estimates of $\mathcal R$ based on the serial- and generation-interval distributions. 
To do so, we model the intrinsic generation-interval distribution and the incubation period using a multivariate log-normal distribution with log means $\mu_G, \mu_I$, log standard variances $\sigma_G^2, \sigma_I^2$, and correlation $\rho$;
the multivariate log-normal distribution is parameterized based on pameter estimates for COVID-19 (Table 1).
We construct forward serial intervals during the exponential growth period as follows:
\begin{equation}
S_i = -B_i + (G_i|B_i) + I_i,
\end{equation}
where the backward incubation period $B_i$ of an infector is simulated by drawing random log-normal samples $A_i$ with log mean $\mu_I$ and log variance $\sigma_I^2$ and resampling $A_i$, each weighted by the inverse of the exponential growth function $\exp(-rA_i)$;
the intrinsic generation interval conditional on the incubation period of the infector $(G_i|B_i)$ is drawn from a log-normal distribution with log mean $\mu_G + \sigma_G \rho (\log(B_i) - \mu_I)/\sigma_I$ and log variance $\sigma_G^2 (1-\rho^2)$;
the forward incubation period $I_i$ of an infectee is drawn from a log-normal distribution with log mean $\mu_I$ and log variance $\sigma_I^2$.
We then calculate the reproduction number $\mathcal R$ using the empirical estimator:
\begin{equation}
\mathcal R = \frac{1}{\frac{1}{N}\sum_{i=1}^N \exp(- r S_i)}.
\end{equation}
We compare this with an estimate of $\mathcal R$ based on naive serial-interval distribution that assumes that the backward and the forward incubation periods are identically distributed \citep{svensson2007note,klinkenberg2011correlation,champredon2018equivalence, britton2019estimation}:
\begin{equation}
{\mathcal R}_{\textrm{\tiny naive}} = \frac{1}{\frac{1}{N}\sum_{i=1}^N \exp(- r T_i)},
\end{equation}
where
\begin{equation}
T_i = -A_i + (G_i|A_i) + I_i.
\end{equation}

@@

First, we compare the estimates of the reproduction number $\mathcal R$ based on the intrinsic generation-interval distribution $g(\tau)$ and the forward serial-interval distribution during the exponential growth phase $f_{\textrm{\tiny exp}}(\tau)$.
Both distributions provide identical estimates of $\mathcal R$ regardless of the correlation $\rho$ between the incubation period distribution and the intrinsic generation-interval distribution (\fref{rR}A).
On the other hand, using naive serial-interval distributions that do not account for disease dynamics (i.e., assuming that the backward and the forward incubation period distributions are identical) underestimates $\mathcal R$;
as $r$ increases, ${\mathcal R}_{\textrm{\tiny naive}}$ saturates and eventually decreases due to negative serial intervals (\fref{rR}B).
While the forward serial intervals during the exponential growth phase can be also negative, the proportion of negative intervals are appropriately balanced because faster epidemic growth will lead to shorter backward incubation period (and therefore less negative serial intervals).

@@

Comapring the shapes of forward serial-interval distributions and the intrinsic generation-interval distribution allow us to better understand how they are able to give identical estimates of $\mathcal R$.
In general, generation-interval distributions with higher means and less variability are expected to give higher $\mathcal R$ for a given $r$.
In this case, the forward serial intervals during the exponential growth phase have higher means (\fref{rR}C) and squared coefficients of variation (\fref{rR}D) than the intrinsic generation-interval distribution.
The effects of higher means (which increases $\mathcal R$) and higher variability (which decreases $\mathcal R$) cancel out exactly;
therefore, we estimate the same $\mathcal R$ using both serial and generation intervals.

@@

\fref{epi} compares the epidemiological dynamics (A) with the mean forward (B--D) and the mean backward (E--F) delay distributions of a deterministic model based on the renewal equation (\eref{renewal}) and the corresponding stochastic realizations based on the Gillespie algorithm.
The mean forward incubation period remains constant throughout an epidemic as expected (\fref{epi}B).
The mean forward generation interval contracts as the epidemic progresses because an infected individual is less likely to infect another person as the proportion of susceptible individuals decreases (\fref{epi}C; \cite{champredon2015intrinsic}).
In contrast, the mean serial interval depends on previous incidence and therefore decreases over time (\fref{epi}D):
When incidence is increasing, symptomatic individuals are more likely to have been infected more recently (shorter backward incubation period and therefore longer forward serial interval) whereas when incidence is decreasing, symptomatic individuals are more likely to have been infected later (longer backward incubation period and therefore shorter forward serial interval).
Qualitative patterns in the changes in the mean backward delays is robust across all delay distributions because they are predominantly driven by the changes in incidence (\fref{epi}D).

\begin{figure}[!th]
\includegraphics[width=\textwidth]{forward_tease.pdf}
\caption{
\textbf{Dynamical correlations between incubation periods and the number of secondary cases.}
A. The mean forward serial interval depends on changes in the cohort reproduction number.
B. The mean backward incubation period of infectors is higher than the mean backward incubation period of all symptomatic cases.
C. Correlation between the number of secondary cases and the incubation period for individuals who got infected within the same time preiod.
D. Correlation between the number of secondary cases and the incubation period for individuals who developed symptoms within the same time preiod.
Black lines represent the results of a deterministic simulation.
Orange lines and points represent the average of 10 stochastic simulations.
Orange ribbons represent the 95\% confidence intervals.
}
\label{fig:tease}
\end{figure}

We further tease apart factors that affect the forward serial interval distribution.
\fref{tease}A compares the mean forward serial interval calculated from a deterministic model and its corresponding stochastic simulations;
failing to account for changes in $\mathcal R_c$ in \eref{forward} overestimates the mean forward serial interval.
A key distinction that needs to be made is that the forward serial-interval distributions depend on the backward incubation period of symptomatic infectors (i.e., those who were able to successfully spread the disease to another person), rather than the incubation period of all symptomatic cases (including those who were not able to spread the disease).
Therefore, the cohort size of infectors must be calculated as a product of the cohort reproduction number and incidence because cohorts with higher cohort reproduction numbers are more infectiouss and therefore more likely to be infectors.

Comparing the backward incubation period of infectors and all symptomatic cases provides an important, but counterintuitive, insight (\fref{tease}B).
Given a cohort of symptomatic cases, individuals who were able to successfully spread the disease to others have longer incubation periods, on average,
because individuals with longer incubation period are infected earlier and have a higher chance of infecting others (i.e., higher $\mathcal R_c$) during the susceptible depletion phase.
Therefore, the backward incubation period of a symptomatic case and the number of secondary cases caused by that individual are dynamically correlated (\fref{tease}D).
This correlation disappears when we take the forward perspective, demonstrating the importance of perspective-dependency in studying epidemiological delay distributions (\fref{tease}C).

>>>>

Here, we provide an answer to this paradox, by showing that the relevant interval for the renewal framework is what is called the ``forward'' interval, and that the forward serial (but not generation) interval is affected by the rate of growth $r$ during the exponential-growth phase.
We develop a new framework for understanding serial intervals and show that the initial forward serial-interval distribution gives the correct value of $\mathcal R$.
%\jd{Come back to this. Not just over time, right?} \swp{What do you mean? Everything is about time here... you haven't read the results section so maybe that's why?}
However, using inaccurately defined serial intervals or failing to account for changes in the observed serial-interval distributions over the course of an epidemic can bias the estimate of $\mathcal R$.
We apply our framework to serial intervals of COVID-19 and lay out several principles to consider in using information about serial intervals and other epidemiological time delays in the analysis of the ongoing pandemic.

@@

In order to define forward and backward serial interval distributions, we begin by writing the total number of serial intervals between time $\pi$ (when the infector developed symptoms) and $\delta$ (when the infectee developed symptoms), which can be expressed as an integral over all of the possibilities for the infection times of the infector $\alpha_1$ and the infectee $\alpha_2$:
\begin{equation}
T(\pi,\delta) = \int_{-\infty}^{\pi} \int_{\alpha_1}^{\delta} \mathcal R_c (\alpha_1) i(\alpha_1) h_{\alpha_1}(\pi-\alpha_1, \alpha_2 - \alpha_1) k_{\alpha_2}(\delta - \alpha_2) \mathrm{d}\alpha_2\,\mathrm{d}\alpha_1,
\end{equation}
where $\mathcal R_c(\alpha_1)$ is the case reproduction number (i.e., the average number of secondary cases caused by a primary case infected at time $\alpha_1$, \cite{fraser2007estimating}), $i(\alpha_1)$ is incidence, $h_{\alpha_1}(\pi-\alpha_1, \alpha_2 - \alpha_1)$ is the joint probability distribution describing the forward incubation period $\pi-\alpha_1$ and forward generation interval $\alpha_2 - \alpha_1$ of an individual infected at time $\alpha_1$ (in this case, the infector), and $k_{\alpha_2}(\delta-\alpha_2)$ describes the marginal probability distribution of $h_{\alpha_2}$ describing the forward incubation period $\delta-\alpha_2$ of an individual infected at time $\alpha_2$ (in this case, the infectee). 
Then, the forward serial-interval distribution is proportional to $T(\pi, \pi+\tau)$. Substituting $x=\pi-\alpha_1$ and $\sigma=\alpha_2-\alpha_1$, we have:
\begin{equation}
f_\pi(\tau) \propto \int_{0}^{\infty} \int_{0}^{\max(0,x+\tau)} \mathcal R_c (\pi-x) i(\pi-x) h_{\pi-x}(x, \sigma) k_{\pi-x+\sigma}(x-\sigma+\tau) \mathrm{d}\sigma\,\mathrm{d}x
\end{equation}
Likewise, the backward serial-interval distribution is proportional to $T(\delta-\tau, \delta)$. 
This time, we substitute $x=\delta-\alpha_2$ and $\sigma=\alpha_2-\alpha_1$ to get:
\begin{equation}
b_\delta(\tau) \propto \int_{0}^{\infty} \int_{\max(0, \tau-x)}^{\infty} \mathcal R_c (\delta-x-\sigma) i(\delta-x-\sigma) h_{\delta-x-\sigma}(x+\sigma-\tau, \sigma) k_{\delta-x}(x) \mathrm{d}\sigma\,\mathrm{d}x.
\end{equation}

@@

We validate the theory by applying it to a specific example of an epidemic model. 
We model disease spread with a renewal-equation model \citep{heesterbeek1996concept, diekmann2000mathematical, roberts2004modelling, aldis2005integral, roberts2007model, champredon2018equivalence}.
Ignoring births and deaths, changes in the proportion of susceptible individuals $S(t)$ and incidence of infection $i(t)$ can be written as:
\begin{equation}
\begin{aligned}
\frac{\mathrm{d}S}{\mathrm{d}t} &= - i(t)\\
i(t) &= \mathcal R_0 S(t) \int_0^\infty i(t-\tau) g(\tau) \mathrm{d}\tau,
\end{aligned}
\label{eq:renewal}
\end{equation}
where $\mathcal R_0$ is the basic reproduction number, and $g(\tau)$ is the intrinsic generation-interval distribution (i.e., the forward generation-interval distribution of a primary case in a fully susceptible population; \cite{champredon2015intrinsic}).
Then, the forward generation-interval for a primary cohort $\pi$ follows \citep{champredon2015intrinsic}:
\begin{equation}
g_\pi (\tau) \propto g(\tau) S(\pi + \tau),
\end{equation}
which allows us to separate the joint probability distribution $h_\pi$ of the forward incubation period and the forward generation-interval distribution as a product of the proportion of susceptible individuals $S$ and the joint probability distribution $h$ of the forward incubation period and the intrinsic generation intervals:
\begin{equation}
h_\pi (x, \tau) \propto h(x, \tau) S(\pi + \tau).
\end{equation}
We further assume that the forward incubation period distributions does not vary across cohorts over the course of an epidemic, as they represent the natural history of a disease, and use $k$ without subscript instead. 
Then, we have:
\begin{equation}
\begin{aligned}
k(x) &= \int_0^\infty h(x, \tau) \mathrm{d}\tau\\
g(\tau) &= \int_0^\infty h(x, \tau) \mathrm{d}x.
\end{aligned}
\end{equation}
Finally, the case reproduction for this model is defined as follows:
\begin{equation}
\mathcal R_c(t) = \mathcal R_0 \int_0^\infty g(\tau) S(t+\tau) \mathrm{d} \tau.
\end{equation}

@@

During the initial phase of an epidemic, the proprotion susceptible remains constant ($S(t) = S(0)$) and incidence of infection grows exponentially: $i(t)=i_0\exp(rt)$.
Then, we can estimate the reproduction number from the exponential growth rate $r$ via the Euler-Lotka equation:
\begin{equation}
\frac{1}{\mathcal R} = \int_0^\infty \exp(-r\tau) g(\tau) \mathrm{d} \tau.
\end{equation}
Like forward generation-interval distributions, 
forward serial-interval distributions describe the renewal process of symptomatic cases.
Therefore, we expect the forward serial-interval distribution during the exponential growth phase --- which we refer to as the \emph{initial} forward serial-interval distribution $f_0$ --- to provide the identical $r$--$\mathcal R$ link as the intrinsic generation-interval distribution:
\begin{equation}
\frac{1}{\mathcal R} = \int_{-\infty}^\infty \exp(-r\tau) f_{0}(\tau) \mathrm{d} \tau,
\end{equation}
where the initial forward serial-interval distribution is defined as:
\begin{equation}
f_{0}(\tau) \propto \int_{0}^{\infty} \int_{0}^{\max(0,x+\tau)} \exp(-rx) h(x, \sigma) k(x-\sigma+\tau) \mathrm{d}\sigma\,\mathrm{d}x
\end{equation}
In Appendix, we provide a mathematical proof that this relationship holds.

The initial forward serial-interval distribution depends on the exponential growth rate $r$.
For a fast-growing epidemic (high $r$), we expect the backward incubation periods to be short, and therefore, the forward serial-interval distribution will generally have a larger mean than the intrinsic generation-interval distribution.
The Susceptible-Exposed-Infected-Recovered model, which assumes that incubation and exposed periods are equivalent, is a special case where the conditional forward generation-interval distribution cancels out with the backward incubation period distribution exactly because (i) infected individuals can only transmit after symptom onset and (ii) the time between symptom onset to infection is independent of the incubation period of an infector;
in this case, the forward serial- and generation-intervals follow the same distributions during the exponential growth phase.

@@

We use a simulation-based approach to compare the estimates of $\mathcal R$ based on the serial- and generation-interval distributions. 
To do so, we model the intrinsic generation-interval distribution and the incubation period using a multivariate log-normal distribution with log means $\mu_G, \mu_I$, log standard variances $\sigma_G^2, \sigma_I^2$, and correlation $\rho$;
the multivariate log-normal distribution is parameterized based on pameter estimates for COVID-19 (Table 1).
We construct forward serial intervals during the exponential growth period as follows:
\begin{equation}
S_i = -B_i + (G_i|B_i) + I_i,
\end{equation}
where the backward incubation period $B_i$ of an infector is simulated by drawing random log-normal samples $A_i$ with log mean $\mu_I$ and log variance $\sigma_I^2$ and resampling $A_i$, each weighted by the inverse of the exponential growth function $\exp(-rA_i)$;
the intrinsic generation interval conditional on the incubation period of the infector $(G_i|B_i)$ is drawn from a log-normal distribution with log mean $\mu_G + \sigma_G \rho (\log(B_i) - \mu_I)/\sigma_I$ and log variance $\sigma_G^2 (1-\rho^2)$;
the forward incubation period $I_i$ of an infectee is drawn from a log-normal distribution with log mean $\mu_I$ and log variance $\sigma_I^2$.
We then calculate the reproduction number $\mathcal R$ using the empirical estimator:
\begin{equation}
\mathcal R = \frac{1}{\frac{1}{N}\sum_{i=1}^N \exp(- r S_i)}.
\end{equation}
We compare this with an estimate of $\mathcal R$ based on naive serial-interval distribution that assumes that the backward and the forward incubation periods are identically distributed \citep{svensson2007note,klinkenberg2011correlation,champredon2018equivalence, britton2019estimation}:
\begin{equation}
{\mathcal R}_{\textrm{\tiny naive}} = \frac{1}{\frac{1}{N}\sum_{i=1}^N \exp(- r Q_i)},
\end{equation}
where
\begin{equation}
Q_i = -A_i + (G_i|A_i) + I_i.
\end{equation}

@@

The initial forward serial-interval distributions $f_0(\tau)$ and the intrinsic generation-interval distribution $g(\tau)$ give identical estimates of $\mathcal R$ regardless of the correlation $\rho$ between the incubation period distribution and the intrinsic generation-interval distribution (\fref{rR}A).
However, using naive serial-interval distributions that do not account for disease dynamics (i.e., assuming that the backward and the forward incubation period distributions are identical) underestimates $\mathcal R$;
as $r$ increases, ${\mathcal R}_{\textrm{\tiny naive}}$ saturates and eventually decreases due to negative serial intervals (\fref{rR}B).
While the forward serial intervals during the exponential growth phase can be also negative, the proportion of negative intervals are appropriately balanced because faster epidemic growth will lead to shorter backward incubation periods (and therefore a lower proportion of negative serial intervals).

@@

Comapring the shapes of initial forward serial-interval distributions and the intrinsic generation-interval distribution allows us to better understand how different distributoins are able to give identical estimates of $\mathcal R$.
In general, generation-interval distributions with higher means and less variability are expected to give higher $\mathcal R$ for a given $r$.
In this case, the initial forward serial intervals during the exponential growth phase have higher means (\fref{rR}C) and squared coefficients of variation (\fref{rR}D) than the intrinsic generation-interval distribution.
The effects of higher means (which increases $\mathcal R$) and higher variability (which decreases $\mathcal R$) cancel out exactly;
therefore, we can estimate the same $\mathcal R$ using both serial and generation intervals.

@@

\fref{epi} compares the epidemiological dynamics (A) with the mean forward (B--D) and the mean backward (E--F) delay distributions of a deterministic model based on the renewal equation (\eref{renewal}) and of the corresponding stochastic realizations based on the Gillespie algorithm.
The mean forward incubation period remains constant throughout an epidemic as assumed (\fref{epi}B).
The mean forward generation interval decreases slightly as the epidemic progresses because an infected individual is less likely to infect another person as the proportion of susceptible individuals decreases (\fref{epi}C; \cite{kenah2008generation, champredon2015intrinsic}).
In contrast, the mean forward serial interval decreases over time (\fref{epi}D).

Earlier, we demonstrated that the forward serial interval distributions depend on three intervals (\fref{diagram}): (i) the backward incubation periods of infectors, (ii) the forward generation intervals, and (iii) the forward incubation periods of infectees.
Since both forward incubation period (\fref{epi}B) and generation-interval (\fref{epi}C) distributions remain roughly constant, changes in the forward serial-interval distributions are predominantly driven by the changes in the backward incubation period distributions, whose mean increases over time (\fref{epi}D).
In this case, the increase in the mean backward incubation period directly translates to the decrease in the mean forward serial interval.

Qualitative patterns in the changes in the mean backward delays is robust across all delay distributions because they are predominantly driven by the changes in incidence (\fref{epi}D--F).
During the exponential growth phase, the backward delays are short: individuals are
